Improving Estimates of the State of Global Fisheries <br> Depends on Better Data

### 2021-06-23 documents/ovando-etal-assessing-global-fisheries.Rmd
jim edits and upating equation numbering

---
title: Improving Estimates of the State of Global Fisheries <br> Depends on Better Data
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: fish-and-fisheries.csl
output: 
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
always_allow_html: true
linkcolor: blue
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
toc: false
---

@@

**Authors**:
Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^

@@

\begin{equation}
  p_t \sim e^{N\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
(\#eq:procerror)
\end{equation}

@@

When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in Stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018]. By default the model uses 2000 draws with a 1000 step warm-up and one chain. Any detailed fit for an individual fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria. The model fits to a direct estimate of abundance (e.g. a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood is

$$log(a_t) \sim N(f_{pt}(r,K,m,B_0,\pmb{p},\pmb{c}) \times q, \sigma_{obs})$$

where $a_t$ is the observed abundance index and $f_{pt}$ is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)). When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data. @rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep). Treating this raw effort data as the denominator in the CPUE calculation assumes that every increase in fishing effort translates to a commensurate increase in fishing mortality. When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population. This is due to the fact that in reality the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the realized fishing mortality rate must be between between 0% and 100%). To accommodate this, we generate a catch per effective harvest rate index of abundance, as

$$cpue_t = \frac{c_t}{(1 - e^{-F_t})}$$

$$F_t = q_tE_t$$

@@

$$q_t = q_{t-1} \times (1  + \tau)$$

@@

$$log(cpue_t) \sim N(f_{pt}(r,K,m,B_0,\pmb{p},\pmb{c}), \sigma_{obs})$$

@@

$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$

@@

$$p(n_i) = \frac{p(bin_i)}{N_i}$$

@@

Our claim is that achieving substantial progress in assessing the state of global fisheries will require improved data and capacity building. This claim cannot be definitively proven; it is possible that some yet-to-be discovered model will provide a reliable means of dramatically improving the accuracy of estimates produced by catch-based models using current broadly-available datasets. Basic logic tells us though that such a model would have to depend on strong assumptions, new data, or empirical estimation. Catch is a function of catchability, effort, and biomass, so given just catch we have one equation and three unknowns. Structurally separating out biomass (and subsequently estimating reference points etc.) requires then either data or strong assumptions on these other variables across a sufficiently representative timespan. Alternatively, empirical relationships could be developed linking attributes of catch histories to observed estimates of stock status [e.g. @costello2012b]. However, these methods require that these empirical relationships both exist and are reliably preserved in fisheries that lack formal assessments. Numerous groups made up of highly skilled scientists have attempted to resolve these fundamental challenges in various ways, but external efforts to validate the performance of these methods have consistently found them to perform generally poorly when applied in bulk [@free2020;@bouch2021;@pons2020]. This is not intended as an indictment of those past efforts, but, we believe, a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone. Methods tested in those papers have been demonstrated to perform relatively well when given specific data or priors informative to specific fisheries. We revisited this task in this paper, and found that even when augmented with other broadly available datasets, we were unable to produce consistently accurate estimates of stock status around the globe. For these reasons we believe that improvements to our understanding of the state of global unassessed fisheries will come from targeted use of existing but underutilized data, and expanding the collection of high-priority data around the globe, rather than the development of new modeling methods alone.

>>>>

---
title: Improving Estimates of the State of Global Fisheries <br> Depends on Better Data
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: fish-and-fisheries.csl
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
  bookdown::word_document2:
    reference_docx: template.docx
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
always_allow_html: true
linkcolor: blue
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
toc: false
---

@@

**Authors**:
Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James T. Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^

@@

\begin{equation}
  log(p_t) \sim {N}\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)
(\#eq:procerror)
\end{equation}

@@

When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in Stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018]. By default the model uses 2000 draws with a 1000 step warm-up and one chain. Any detailed fit for an individual fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria. The model fits to a direct estimate of abundance (e.g. a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood calculated per

\begin{equation}
  B_t = f_{t}(r,K,m,B_0,\pmb{p},\pmb{c})
(\#eq:biomass)
\end{equation}

where $B_t$ is the biomass in time *t*, given $f_{t}$ which is shorthand for the result of the Pella-Tomlinson equation in time step *t* given parameters and a vector of catches. The likelihood of the abundance index *a* in time *t* is then calculated as

\begin{equation}
 log(a_t) \sim N(B_t \times q, \sigma_{obs})
(\#eq:likelihood)
\end{equation}

When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data. @rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep). Treating this raw effort data as the denominator in the CPUE calculation assumes that every increase in fishing effort translates to a commensurate increase in fishing mortality. When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population. This is due to the fact that in reality the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the realized exploitation fraction, sometimes called *u*, is between 0 and 1). To accommodate this, we generate a catch per effective harvest rate index of abundance, as

\begin{equation}
 cpue_t = \frac{c_t}{(1 - e^{-F_t})}
(\#eq:cpue)
\end{equation}

\begin{equation}
 F_t = q_tE_t
(\#eq:f)
\end{equation}

@@

\begin{equation}
 q_t = q_{t-1} \times (1  + \tau)
(\#eq:qt)
\end{equation}

@@

\begin{equation}
 log(cpue_t) \sim N(f_{pt}(r,K,m,B_0,\pmb{p},\pmb{c}), \sigma_{obs})
(\#eq:cpue)
\end{equation}

@@

\begin{equation}
 p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}
(\#eq:pbin)
\end{equation}

@@

\begin{equation}
 p(n_i) = \frac{p(bin_i)}{N_i}
(\#eq:ppcor)
\end{equation}

@@

Our claim is that achieving substantial progress in assessing the state of global fisheries will require improved data and capacity building. This claim cannot be definitively proven; it is possible that some yet-to-be discovered model will provide a reliable means of dramatically improving the accuracy of estimates produced by catch-based models using current broadly-available datasets. Basic logic tells us though that such a model would have to depend on strong assumptions, new data, or empirical estimation. Catch is a function of catchability, effort, and biomass, so given just catch we have one equation and three unknowns. Structurally separating out biomass (and subsequently estimating reference points etc.) requires then either data or strong assumptions on these other variables across a sufficiently representative timespan. Alternatively, empirical relationships could be developed linking attributes of catch histories to observed estimates of stock status [@costello2012b;@thorson2012a], or through identifying a structural model for the evolution of fishing mortality in managed and unmanaged fisheries [@thorson2013]. However, these methods require that these empirical relationships both exist and are reliably preserved in fisheries that lack formal assessments. Numerous groups made up of highly skilled scientists have attempted to resolve these fundamental challenges in various ways, but external efforts to validate the performance of these methods have consistently found them to perform generally poorly when applied in bulk [@free2020;@bouch2021;@pons2020]. This is not intended as an indictment of those past efforts, but, we believe, a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone. Methods tested in those papers have been demonstrated to perform relatively well when given specific data or priors informative to specific fisheries. We revisited this task in this paper, and found that even when augmented with other broadly available datasets, we were unable to produce consistently accurate estimates of stock status around the globe. For these reasons we believe that improvements to our understanding of the state of global unassessed fisheries will come from targeted use of existing but underutilized data, and expanding the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
